---
title: "Harry Potter and the Art of Fanfiction: Data Cleaning"
author: "Isabel Leiva"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 2
    css: !expr here::here("../StylesTemplates/il_hp_26.css")
---
```{r Libraries, include = F}
#wipe R's memory
rm(list=ls())

#condensed way of writing library(xyz package)
isa_packages <- c("plyr", "dplyr", "tidyr", "rlang", "readr", "ggplot2", "GGally",  
              "skimr", "lme4", "lmerTest", "effects", "reshape", "reshape2",
              "psych", "tidyverse", "modelsummary", "lme4", "sjPlot", "rstatix",
              "glmnet", "apaTables","umx", "ltm","effectsize","emmeans","ggeffects",
              "MuMIn","stats", "here", "kableExtra", "RCurl",  "knitr", "RColorBrewer",
              'gutenbergr', 'janeaustenr',"tidytext","xfun","tm", "stringr", "stringi", 
              "quanteda", "quanteda.textplots", "textstem","quanteda.textstats",
              "textclean", "readtext", "udpipe", "harrypotter", "praise", "devtools")

packages <- rownames(installed.packages())
p_to_install <- isa_packages[!(isa_packages %in% packages)]

if(length(p_to_install) > 0){
  install.packages(p_to_install)
}

lapply(isa_packages, library, character.only = TRUE)
```

```{r knitr_setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2

#prints the knitted output with these specifications
print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Blippo", font_size = 12) %>%
   row_spec(0, color = "white", background = "#745AAB", font_size = 12, bold = TRUE) %>%
   #scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#to_R <- read.csv(here("data", "MyRaw.txt")) ---- using 'here' to read in data
```

# Research Question

It is often said that fanfiction is "darker" thematically compared to the original works they are based off of. Some even go as far as to say that these more explicit themes is what makes the fanfiction themselves original works. However, this begs the question of whether fanfiction is "darker" than the original works in some tangible way. Thankfully, with the advent of natural language processing and semantic analysis, we can finally answer this afe old question: 

Does fanfiction differ significantly in overall negative valence and negative valence over time from the works that it draws from?

We will be testing this using one of the most popular fandoms of the 21st century, Harry Potter.

# Data Preperation

All of the original texts are available in [Bradley Boemke's 'harrypotter' package](https://github.com/bradleyboehmke/harrypotterpackage). <br /> 

Note: there are two 'harrypotter' packages. To snag the correct one, use code snippet: devtools::install_github("bradleyboehmke/harrypotter") <br />

All of the Fanfiction is available on Archive of our own [https://archiveofourown.org/]
<br />

## Original Harry Potter Cleanup

```{r og_hp_clean}
#downloading the harry potter novels from the package
#devtools::install_github("bradleyboehmke/harrypotter")
data(package = "harrypotter")
data(philosophers_stone)
data(chamber_of_secrets)
data(prisoner_of_azkaban)
data(goblet_of_fire)
data(order_of_the_phoenix)
data(half_blood_prince)
data(deathly_hallows)

#collapsing the novels into one column and row
p <- philosophers_stone %>% #write to an object
  paste(collapse= " ") #collapse character vector
c <- chamber_of_secrets %>% #write to an object
  paste(collapse= " ") #collapse character vector
a <- prisoner_of_azkaban %>% #write to an object
  paste(collapse= " ") #collapse character vector
g <- goblet_of_fire %>% #write to an object
  paste(collapse= " ") #collapse character vector
o <- order_of_the_phoenix %>% #write to an object
  paste(collapse= " ") #collapse character vector
h <- half_blood_prince %>% #write to an object
  paste(collapse= " ") #collapse character vector
d <- deathly_hallows %>% #write to an object
  paste(collapse= " ") #collapse character vector

#creating dataframe with books text
hp_og <- data.frame(text = rbind(p, c, a, g, o, h, d))

#creating new author, genre, & doc_id
hp_og <- hp_og %>% 
  mutate(author = c(1, 1, 1, 1, 1, 1, 1),
         genre = c("og", "og", "og", "og", "og", "og", "og"),
         doc_id = c("philosophers_stone", "chamber_of_secrets", "prisoner_of_azkaban", "goblet_of_fire", "order_of_the_phoenix", "half_blood_prince", "deathly_hallows"))
```

## Harry Potter Fanfic Cleanup

```{r hp_fanfic_clean}
#creating a df with all IDs
#Directory containing CSV files
fic_directory <- "fanfics"

#List of .txt files in the directory
txt_files <- list.files(path = fic_directory, pattern = "*.txt", full.names = TRUE)

#reading in text files as dataframes
dataframes <- lapply(txt_files, readtext)

# rbind all dataframes
hp_fanfic <- do.call(rbind, dataframes)

#creating author column
hp_fanfic$author <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 8, 14, 15, 16, 17, 18, 19, 20) #choices and kill your darlings have the same author

#creating genre column
hp_fanfic$genre <- c("fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic", "fanfic")

```

## Joining Dataframes

```{r joining_dataframes}
all_hp <- rbind(hp_og, hp_fanfic)
all_hp <- all_hp %>% dplyr::select(doc_id, genre, author, text)
rownames(all_hp) <- NULL

#making the following variables factors
all_hp$doc_id <- as.factor(all_hp$doc_id)
all_hp$genre <-  as.factor(all_hp$genre)
all_hp$author <-  as.factor(all_hp$author)

#confirming variables are now factors
str(all_hp$doc_id)
str(all_hp$genre)
str(all_hp$author)
```

## Loading & cleaning sentiment database

```{r sentiment_database}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))
lookup_db <- lookup_Jul25 #renaming dataframe
colnames(lookup_db)
lookup_small <- lookup_db %>% dplyr::select(word, emo_valence_b24_rescale,emo_sadness_rescale) #creating database that suites hypotheses being tested
lookup_small[1:20,] #printing the first 20 lines of the df
lookup_small[lookup_small$word == "smile",] #checking directionality
lookup_small[lookup_small$word == "cry",] #checking directionality

```

## Cleaning genie

```{r cleaning_genie, eval=T}
cleaning_genie <- function(dat, wordcol, omit_stops = TRUE, lemmatize = TRUE) {
  # Input validation
  #telling R what column in your dataframe is the text in quotes 'wordcol', this error message checks
  if (!wordcol %in% names(dat)) {
    stop(paste("Column", wordcol, "not found in dataframe"))
  }
#load stopword and replacement lists from github  
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))

# Creating a working df and standardizing
  dat_prep <- dat %>% #creating a working df called "dat prep"
    dplyr::mutate(id_row_orig = factor(seq_len(nrow(dat))), # creating a new row that tracks id number (starting from 1)
      text_initialsplit = tryCatch(stringi::stri_enc_toutf8(as.character(.[[wordcol]]), #creating new row that standardizes text as UTF8
                                 is_unknown_8bit = TRUE,
                                 validate = TRUE), error = function(e) stringi::stri_encode(as.character(.[[wordcol]]), to = "UTF-8")
      ) %>% tolower()) #makes all text lower case

  # Standardize apostrophes in original text
  dat_prep <- dat_prep %>% dplyr::mutate(text_initialsplit = ifelse(
        is.na(text_initialsplit), NA_character_, stringi::stri_replace_all_regex(
          text_initialsplit,
          "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]",
          "'"
        )
      )
    )

  # Perform initial split into words (retains contractions)
  dat_prep <- dat_prep %>% tidyr::separate_rows(text_initialsplit, sep = "[[:space:]]+") %>%
    dplyr::mutate(
      text_initialsplit = ifelse(
        is.na(text_initialsplit) | stringi::stri_isempty(text_initialsplit),
        NA_character_,
        text_initialsplit
      )
    ) %>%
    # Remove original column
    dplyr::select(-all_of(wordcol))

  # Creating clean column (called "word_clean") with actual cleaned text
  dat_prep <- dat_prep %>%
    dplyr::mutate(
      word_clean = text_initialsplit,
      # Remove non-alphabetic characters except apostrophes
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-zA-Z']", " ")
      ),
      # Remove whitespace
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringr::str_squish(word_clean)
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-z']", "")
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^[:alnum:]']", "")
      )
    )

#Replace words function - this finds contractions and replaces them with their expanded forms from a lookup database. 
 replace_words <- function(word_clean, replacement_df) {
 for (i in seq_len(nrow(replacement_df))) {
   text_vector <- gsub(
     pattern = replacement_df$word[i],
     replacement = replacement_df$replacement[i],
     x = word_clean, ignore.case = TRUE)
 }
 return(word_clean)
 }
 
 #run replace words function on dat_prep
dat_prep <- replace_words(dat_prep, replacement_df = replacements_25)

  # Perform additional splitting after replacements (keeping NAs)
  dat_prep <- dat_prep %>%
    tidyr::separate_rows(word_clean, sep = "[[:space:]]+", convert = TRUE) %>%
    dplyr::mutate(
      word_clean = ifelse(
        is.na(word_clean) | stringi::stri_isempty(word_clean),
        NA_character_,
        word_clean
      ),
      id_row_postsplit = seq_len(dplyr::n())
    )

  # Lemmatization if requested (only for non-NA values)
  if (lemmatize) {
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        word_clean = ifelse(
          is.na(word_clean),
          NA_character_,
          textstem::lemmatize_strings(word_clean)
        )
      )
  }

  # Stopword removal if requested (now keeps NAs)
  if (omit_stops) {
    stopwords <- tolower(Temple_stops25$word)
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        is_stopword = ifelse(
          is.na(word_clean),
          NA,
          word_clean %in% stopwords
        ),
        word_clean = ifelse(is_stopword, NA_character_, word_clean)
      ) %>%
      dplyr::select(-is_stopword)  # Remove the is_stopword column
  }

  return(dat_prep)
}

```

## Deploy Genie

```{r using_genie, eval=F}
hp_clean <- cleaning_genie(all_hp, 'text')

write.csv(hp_clean, file = "hp_clean.csv")
```
