---
title: "Homework 3"
author: "Isabel Leiva"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 2
    css: !expr here::here("../StylesTemplates/il_26.css")
---
```{r Libraries, include = F}
#wipe R's memory
rm(list=ls())

#condensed way of writing library(xyz package)
isa_packages <- c("plyr", "dplyr", "tidyr", "rlang", "readr", "ggplot2", "GGally",  
              "skimr", "lme4", "lmerTest", "effects", "reshape", "reshape2",
              "psych", "tidyverse", "modelsummary", "lme4", "sjPlot", "rstatix",
              "glmnet", "apaTables","umx", "ltm","effectsize","emmeans","ggeffects",
              "MuMIn","stats", "here", "kableExtra", "RCurl",  "knitr", "RColorBrewer",
              'gutenbergr', 'janeaustenr',"tidytext","xfun","tm", "stringi")

packages <- rownames(installed.packages())
p_to_install <- isa_packages[!(isa_packages %in% packages)]

if(length(p_to_install) > 0){
  install.packages(p_to_install)
}

lapply(isa_packages, library, character.only = TRUE)
```

```{r knitr_setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2

#prints the knitted output with these specifications
print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Blippo", font_size = 12) %>%
   row_spec(0, color = "white", background = "#9DC2A1", font_size = 12, bold = TRUE) %>%
   #scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#to_R <- read.csv(here("data", "MyRaw.txt")) ---- using 'here' to read in data
```

# Homework: Creating a "Cleaning Genie" function
1. Remove punctuation
2. Remove numbers
3. Remove "stop words"
4. Lengthen contractions
4. Convert to lower case
5. Remove blank space
6. Split and unlist
7. Remove single characters 

## Loading data
```{r}
load("data/MIT_stops.rda") #MIT stop words
load("data/replacements_25.rda") # contraction replacements
load("data/SMART_stops.rda") #smart stop words
load("data/Temple_stops25.rda") #temple stop words
unabomb <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt')) #unabomber manifesto
cat(unabomb[1:10], sep = "\n") #Print first few lines of unabomber manifesto
```

## Creating functions to streamline the genie
```{r}
# Function to replace a word in a target data frame with a replacement word
replace_words <- function(text, replacements_25) {
 # Replace each bad word with its corresponding good word
 for (i in seq_len(nrow(replacements_25))) {
   text <- gsub(
     pattern = replacements_25$word[i],
     replacement = replacements_25$replacement[i],
     x = text,
     ignore.case = TRUE
   )
 }
 return(text)
}
#note: this ONLY works if the apostrophe's are the same font the apostrophe that my computer creates " ' " is not recognized as matching the apostrophe's on the replacements_25 list, which look like this " ’ ". We may want to discuss that?

```

## Cleaning Genie Function
```{r}
#Cleaning Genie function
cleaning_genie <- function(x){ #stating the function
   y <- tolower(x) #makes all lower case
   y <- replace_words(y, replacements_25) #replaces contractions with full words
   y <- gsub('[[:punct:]]',' ', y) #removing all punctuation
   y <- gsub("\\b[a-zA-Z]\\b{1}", " ", y) #omits singletons including Aa and Ii
   y <- textstem::lemmatize_strings(y) #lemmatizes text
   y <- gsub('[0-9]+',' ', y) #removing all numbers
   y <- tibble(y) #making into a dataframe for the final piece!
   y <- y %>% unnest_tokens(word, y)  # one word per rows, removing blank spaces, and making all lowercase
   y <- y %>% anti_join(SMART_stops) #removing stop words by line
   return(y) #final step of the function
}

```

## Deploy Genie! 
```{r}
unabomb_clean <- cleaning_genie(unabomb) #using the function on unabomber manifesto and saving as new df
unabomb_clean[1:20,] #printing the first 20 lines of the df
```

# Optional Homework (mentioned in class)
1. Make a Document Feature Matrix (DFM) for pride and prejudice

## Pride & prejudice DFM
```{r}
pride_prej <- readLines('data/PridePrejudice.txt')
pride_prej <- data.frame(full_text = pride_prej)
data(stop_words)
pride_prej_clean <- pride_prej %>% #piping through the "pride_prej" df and storing changes in a new df called "sticks_clean"
  mutate(full_text = gsub(pattern = '[[:punct:]]',replacement ='', full_text)) %>% #using mutate to overwrite the "pride_prej" "full_text" col and using gsub to remove all punctuation
  mutate(full_text = gsub(pattern = '[0-9]+',replacement ='', full_text)) %>% #using mutate to overwrite the "pride_prej" "full_text" col and using gsub to remove all numbers
  unnest_tokens(word, full_text) %>% # separating the words into individual rows, removing blank spaces, and making all lowercase
  anti_join(stop_words) #removing stop words
pride_prej_count <- pride_prej_clean %>% count(word, sort = TRUE) #identifying unique words in df and counting total using "length"
pride_prej_count[1:20,] #printing the first 20 most used words
```

